---
title: "Statistikos laboratorinis darbas Nr. 2"
author: "VU"
date: "2025-03-27"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Duomenys atsisiųsti iš <https://github.com/valdas-v1/lithuanian-real-estate-listings>. Duomenys buvo surinkti 2024 m. vasarį iš <https://www.aruodas.lt/> puslapio. Duomenų rinkinyje yra informacija apie parduodamus ir nuomojamus butus, garažus, namus, sklypus ir patalpas.

## Reading files from all folders

```{r}
# Set the path to the data directory
data_dir <- "C:/Users/zabit/Documents/GitHub/Statistikos-lab-2/data"

# Get all folder names inside the data directory
folders <- list.dirs(data_dir, full.names = FALSE, recursive = FALSE)

# Print all folder names
print(folders)

# Check if all folders have the file "all_cities_20240214.csv"
all_have_file <- TRUE
folders_with_file <- 0
folders_missing_file <- character(0)

for (folder in folders) {
  file_path <- file.path(data_dir, folder, "all_cities_20240214.csv")
  if (file.exists(file_path)) {
    folders_with_file <- folders_with_file + 1
  } else {
    all_have_file <- FALSE
    folders_missing_file <- c(folders_missing_file, folder)
  }
}

# Print results
cat("Folders with 'all_cities_20240214.csv':", folders_with_file, "out of", num_folders, "\n")
```

```{r}
# Read the CSV files into a list of dataframes
csv_data_list <- list()

for (folder in folders) {
  file_path <- file.path(data_dir, folder, "all_cities_20240214.csv")
  if (file.exists(file_path)) {
    # Try reading the file
    tryCatch({
      df <- read.csv(file_path)
      csv_data_list[[folder]] <- df
      cat("Read:", folder, "with", nrow(df), "rows and", ncol(df), "columns\n")
    }, error = function(e) {
      cat("Error", folder, ":", conditionMessage(e), "\n")
    })
  }
}
```

```{r}
# First print the number of rows in each dataset
for (type in names(csv_data_list)) {
  if (!is.null(csv_data_list[[type]])) {
    cat(type, ": ", nrow(csv_data_list[[type]]), " rows\n", sep="")
  }
}
```

```{r}
# Remove rows with extreme prices (price > 50,000,000 or price < 20)
for (type in names(csv_data_list)) {
  if (!is.null(csv_data_list[[type]]) && "price" %in% colnames(csv_data_list[[type]])) {
    # Count rows with extreme prices
    extreme_high <- sum(csv_data_list[[type]]$price > 50000000, na.rm = TRUE)
    extreme_low <- sum(csv_data_list[[type]]$price < 20, na.rm = TRUE)
    extreme_prices <- extreme_high + extreme_low
    
    if (extreme_prices > 0) {
      # Store original row count
      original_count <- nrow(csv_data_list[[type]])
      
      # Remove rows with extreme prices, keep rows where price is within range or NA
      csv_data_list[[type]] <- csv_data_list[[type]][(csv_data_list[[type]]$price >= 20 & 
                                                     csv_data_list[[type]]$price <= 50000000) | 
                                                     is.na(csv_data_list[[type]]$price), ]
      
      # Verify how many rows were removed
      new_count <- nrow(csv_data_list[[type]])
      removed_count <- original_count - new_count
      
      cat(type, ": Removed ", removed_count, " rows with extreme prices (", extreme_high, 
          " high, ", extreme_low, " low)\n", sep="")
    } else {
      cat(type, ": No rows with extreme prices\n", sep="")
    }
  } else if (!is.null(csv_data_list[[type]])) {
    cat(type, ": No price column found\n", sep="")
  }
}
```
```{r}
for (type in names(csv_data_list)) {
  if (!is.null(csv_data_list[[type]])) {
    cat(type, ": ", nrow(csv_data_list[[type]]), " rows\n", sep="")
  }
}
```

```{r}
# Output summary of data
cat("\nSummary of all loaded datasets:\n")
for (folder_name in names(csv_data_list)) {
  cat("\nDataset from folder:", folder_name, "\n")
  cat("Number of rows:", nrow(csv_data_list[[folder_name]]), "\n")
  cat("Number of columns:", ncol(csv_data_list[[folder_name]]), "\n")
  cat("Column names:", paste(colnames(csv_data_list[[folder_name]]), collapse = ", "), "\n")
}
```

```{r}
# Real estate types to analyze
real_estate_types <- c("apartments", "houses", "land", "premises")

for (type in real_estate_types) {
  cat("\n----------\n", toupper(type), "PRICE SUMMARY\n")
  
  # Check if this real estate type exists in our list
  if (type %in% names(csv_data_list)) {
    df <- csv_data_list[[type]]
    
    # Check if price column exists
    if ("price" %in% colnames(df)) {
      # Extract price data
      prices <- df$price
      
      # Generate summary statistics
      cat("Number of observations:", length(prices), "\n")
      cat("Summary statistics:\n")
      print(summary(prices))
      
      # Additional statistics
      cat("\nStandard deviation:", sd(prices, na.rm = TRUE), "\n")
    } else {
      cat("No 'price' column found in", type, "dataset.\n")
    }
  } else {
    cat("No data available for", type, "real estate type.\n")
  }
}
```

```{r}
# Load ggplot2 for better visualization
if(!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

# Setting up a 2x2 panel for the plots
par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

# Real estate types to analyze
real_estate_types <- c("apartments", "houses", "land", "premises")

# Loop through each type and create box plots
for (type in real_estate_types) {
  if (type %in% names(csv_data_list) && "price" %in% colnames(csv_data_list[[type]])) {
    # Get price data
    prices <- csv_data_list[[type]]$price
    
    # Basic box plot
    boxplot(prices, main = paste(toupper(type), "PRICES"), 
            ylab = "Price (EUR)", col = "lightblue", 
            outline = TRUE, # Show outliers
            na.rm = TRUE)
    
    # Alternative log-scale box plot (prices often have skewed distributions)
    # Uncomment below if needed
    # boxplot(log10(prices[prices > 0]), main = paste(toupper(type), "PRICES (Log Scale)"),
    #        ylab = "Log10(Price)", col = "lightgreen", outline = TRUE, na.rm = TRUE)
  } else {
    # Create an empty plot with a message if data isn't available
    plot(1, type = "n", xlab = "", ylab = "", 
         main = paste(toupper(type), "- No data available"))
    text(1, 1, "No price data available", col = "red")
  }
}

# Reset the plot layout
par(mfrow = c(1, 1))

# Alternative: Create a more sophisticated plot with ggplot2
# This creates a single plot with all real estate types for comparison
price_data <- data.frame()

for (type in real_estate_types) {
  if (type %in% names(csv_data_list) && "price" %in% colnames(csv_data_list[[type]])) {
    # Extract prices and create a data frame
    temp_data <- data.frame(
      price = csv_data_list[[type]]$price,
      type = rep(type, length(csv_data_list[[type]]$price))
    )
    price_data <- rbind(price_data, temp_data)
  }
}

# Create combined box plot if we have data
if (nrow(price_data) > 0) {
  ggplot(price_data, aes(x = type, y = price, fill = type)) +
    geom_boxplot(outlier.size = 1) +
    scale_y_continuous(labels = scales::comma) +
    labs(title = "Price Distribution Across Real Estate Types",
         x = "Real Estate Type",
         y = "Price (EUR)") +
    theme_minimal() +
    theme(legend.position = "none")
}
```

```{r}

# Reset the plot layout
par(mfrow = c(1, 1))

# Alternative: Create more sophisticated histograms with ggplot2 that use density curves
# This might be better for highly skewed real estate price data
for (type in real_estate_types) {
  if (type %in% names(csv_data_list) && "price" %in% colnames(csv_data_list[[type]])) {
    # Get price data and create a data frame
    df <- data.frame(price = csv_data_list[[type]]$price)
    
    # Create plot object with fixed deprecated features
    p <- ggplot(df, aes(x = price)) +
      geom_histogram(aes(y = after_stat(density)),
                     bins = 30, 
                     fill = "skyblue", 
                     color = "white", 
                     alpha = 0.7) +
      geom_density(color = "darkblue", linewidth = 1) + # Fixed: size -> linewidth
      labs(title = paste(toupper(type), "Price Distribution"),
           x = "Price (EUR)",
           y = "Density") +
      theme_minimal() +
      scale_x_continuous(labels = scales::comma, limits = c(0, 1000000)) +
      coord_cartesian(xlim = c(0, 1000000))
    
    # Print the plot
    print(p)
  }
}


# Create faceted histograms in one figure (comparing all types)
if (nrow(price_data) > 0) {
  ggplot(price_data, aes(x = price, fill = type)) +
    geom_histogram(bins = 30, alpha = 0.7) +
    facet_wrap(~type, scales = "free_y") +
    labs(title = "Price Distributions by Real Estate Type",
         x = "Price (EUR)",
         y = "Count") +
    theme_minimal() +
    scale_x_continuous(labels = scales::comma, limits = c(0, 500000)) +
    coord_cartesian(xlim = c(0, 500000))
}
```
3 užduotis: Kiekybiniai duomenys: kaina ("price"), aukštų skaičius, peržiūrų skaičius, namo, buto dydis ("area")
4. Sudarykite dažnių lenteles kategoriniams kintamiesiems.
5. Suformuluokite bent 6 tyrimo hipotezes iš savo duomenų rinkinio
6. Užrašykite kokius testus parinkote savo tyrimo hipotezėms. Hipotezės turi būti skirtos skirtingų testų
naudojimui. Jei reikia susikurkite naujus kintamuosius iš turimų duomenų.
7. Patikrinkite, ar kintamieji tenkina būtinas sąlygas testų taikymui. Jei netenkina, atlikite duomenų trans-
formacijas.
8. Atlikite statistinį tyrimą savo suformuluotoms hipotezėms.
9. Pateikite tyrimo atsakymą.